---
title: "Julia and R"
output: 
  html_document:
   toc: true
   toc_float: true
   number_sections: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache=TRUE)
```

- [Julia : le language scientifique](https://perso.ensta-paristech.fr/~diam/julia/index.php?id=bm)
- Run tutorials on [juliabox](https://juliabox.com]

## Display notebooks on nbviewer

- [Rosenbrock Julia example](http://nbviewer.jupyter.org/github/StateOfTheR/finistR2018/blob/master/atelier2/julia/JuliaOptim.ipynb)
- [Call julia from R with JuliaCall](http://nbviewer.jupyter.org/github/StateOfTheR/finistR2018/blob/master/atelier2/julia/JuliaCallOptim.ipynb)
- [Call julia from R with XRJulia](http://nbviewer.jupyter.org/github/StateOfTheR/finistR2018/blob/master/atelier2/julia/XRJulia.ipynb)

## Call Julia from R with JuliaCall

[JuliaCall in R Markdown](https://rpubs.com/Consistency/310507)

We check if the package is instyalled or not.

```{r, eval=FALSE}
install.packages("JuliaCall")
```

Then we load it.

```{r, message = FALSE}
library("JuliaCall")
```

We need to inform R about the repository where Julian's binaries are installed

```{r}
julia_setup(JULIA_HOME = "/home/jchiquet/julia-1.0.0/bin")
```

It is then possible to installed Julia package via R: 

```{r}
julia_install_package_if_needed("Optim")
julia_install_package_if_needed("LineSearches")
```

Commands are then executed line by line, which is a bit annoying, but it works !

```{r julia_rosenbrock_optim}
julia_library("Optim")
julia_library("LineSearches")
julia_command("rosenbrock(x) =  (1.0 - x[1])^2 + 100.0 * (x[2] - x[1]^2)^2")
julia_eval("optimize(rosenbrock, zeros(2), BFGS())")
```


## julia chunck

It is also possible to include Julia Chuhnk just like we do with R in Rmarkdown ! First, I needed to link mly Julia's binary to the usual place where Julia is installed and where XRJulia is going to look at: 

```{bash, eval = FALSE}
sudo ln -s /home/jchiquet/julia-1.0.0/bin/julia /usr/local/bin/julia
```

And call direct julia's code: 

```{julia}
using Pkg
Pkg.add("Optim")
Pkg.add("LineSearches")
```

```{julia}
using Optim, LineSearches
```

```{julia}
rosenbrock(x) =  (1.0 - x[1])^2 + 100.0 * (x[2] - x[1]^2)^2
result = optimize(rosenbrock, zeros(2), BFGS())
```

```{julia}
 optimize(rosenbrock, zeros(2), BFGS(linesearch = LineSearches.BackTracking()))
```

```{julia}
optimize(rosenbrock, zeros(2), BFGS(linesearch = LineSearches.BackTracking(order = 2)))
```

For the Rosenbrock example, the analytical gradient can be shown to be:

```{julia}
function g!(G, x)
G[1] = -2.0 * (1.0 - x[1]) - 400.0 * (x[2] - x[1]^2) * x[1]
G[2] = 200.0 * (x[2] - x[1]^2)
end
```

```{julia}
optimize(rosenbrock, g!, zeros(2), LBFGS())
```

In addition to providing gradients, you can provide a Hessian function h! as well. In our current case this is:

```{julia}
function h!(H, x)
    H[1, 1] = 2.0 - 400.0 * x[2] + 1200.0 * x[1]^2
    H[1, 2] = -400.0 * x[1]
    H[2, 1] = -400.0 * x[1]
    H[2, 2] = 200.0
end
```

```{julia}
optimize(rosenbrock, g!, h!, zeros(2))  # newton
```

### References

- [Writing Julia functions in R with examples](https://insightr.wordpress.com/2017/10/18/writing-julia-functions-in-r-with-examples/)
- [XRJulia CRAN](https://cran.r-project.org/web/packages/XRJulia/index.html)
- [XRJulia GitHub](https://github.com/johnmchambers/XRJulia)
